raptr: Representative and Adequate Prioritisation Toolkit in R

### 2016-11-11 code/rmarkdown/article.Rmd
+ further changes based on hugh's comments

4. The \\texttt{raptr R} package provides a unified framework for achieving spatial conservation prioritisations that are both adequate and representative, resulting in a greater chance of achieving long-term biodiversity persistence than area-based planning alone.

@@

<!-- conservation and reserve selection -->
Perhaps the most fundamental aim of conservation is to maximise the long-term persistence of biodiversity [@r423; @r255]. To achieve this, conservation actions must preserve biodiversity patterns (eg. populations, species, ecosystems), but also crucially the processes that sustain them. One of the major tangible achievements of modern conservation has been the act of setting aside areas for preservation [@r440]. Reserve networks buffer species from gross threatening processes and set the stage for enhanced management interventions [@r465]. Since the resources available for conservation action are limited, protected area networks must be sited in places that satisfy conservation objectives for minimal cost [@r255]. However, despite the importance of biodiversity processes, they are only rarely used to guide conservation planning.

<!-- biodiversity processes are important -->
Prioritisations must capture the ecological and evolutionary processes that sustain biodiversity [@r3; @r255; @r9]. Ecological processes are important for biodiversity to persist over short time-scales. For instance, predator-prey processes can mediate ecosystem structure and function through top-down control [@r476]. Additionally, migratory processes allow many species to access suitable habitat throughout the year [@r477]. On the other hand, evolutionary processes are important for biodiversity to persist over long time-scales. Adaptive evolutionary processes foster resilience against environmental change [@r3; @r43], and neutral evolutionary processes maintain genetic diversity [@r5; @r67; @r63]. In recent decades, a wealth of data relevant to biodiversity processes has become available [eg. bioclimatic, distribution, genetic, phylogenetic, and trait data; @r33; @r460; @r478; @r453]. This has sparked significant interest in using these data to make better conservation decisions [@r68; @r9; @r473; @r247; @r448; @r468; @r2; @r407; @r61; @r470; @r479; @r475; @r469].

<!-- current methods for preserving biodiversity processes -->
Today, the most widely used conservation planning tools focus on biodiversity patterns [\texttt{Zonation} and \texttt{Marxan}; @r429; @r22]. These tools allow decision makers to obtain solutions that secure a proportion of the geographic range of biological features (populations, species, or ecosystems of conservation interest). To incorporate some information on biodiversity processes, current methods involve partitioning the features into different groups as a pre-processing step. The features are partitioned based on an attribute variable that relates to the biodiversity process of interest [@r9; @r247; @r448]. For instance, species distributions can be divided according to habitat discontinuities to capture adaptive processes [@r2]. However, using this method is challenging because biodiversity data are often continuous and multi-dimensional.

<!-- issues with partitioning features -->
After partitioning features based on continuous variables and obtaining a solution, it is not clear how well the solution samples the original continuous variables. For instance, if a feature was split into four sub-features based on a single continuous variable, and a solution samples 10 % of each sub-feature, how much of the original continuous variation does the solution sample? This is not immediately clear. If the solution sites reserves evenly along the original variable then it may sample far more than 10 % of the original variation. Whereas, if the solution sites reserves near the values used to divide the continuous values [as observed when partitioning continuous geographic variation based on political boundaries; @r483], then it may sample much less of the original variation. This limitation has not arisen recently. In fact, it dates back to some of the earliest reserve selection methods [eg. @r295].

<!-- diversity method for conservation -->
During the early 1990s, Faith and Walker [-@r480] developed a decision support tool (\texttt{DIVERSITY}) to identify prioritisations that secure a representative sample of continuous environmental variation across a study area. Although this method has been pitched as an alternative to species-based conservation planning [@r481; @r218; @r203], it can also identify solutions that conserve a representative sample of variation across a single species' range. Recent work has built on this reserve selection method by solving problems with more advanced optimisation routines [@r482]. However, \texttt{DIVERSITY}-based reserve selection methods are not widely used today, in part, because they do not accommodate explicit amount-based targets for multiple features (_cf_. \texttt{Marxan} and \texttt{Zonation}).

<!-- aims of this paper -->
One of the key issues in reserve selection is the lack of a unified decision support tool that uses explicit and separate targets to obtain solutions that secure an adequate amount of habitat for each feature and a representative sample of each feature. To begin to fill this gap, here we unite the ideas underpinning \texttt{DIVERSITY} and \texttt{Marxan} into new formulations of the reserve selection problem, and implement them in the \texttt{raptr R} package. This \texttt{R} package provides decision makers with the tools to generate prioritisations based on data that relates to biodiversity patterns and processes. Here, we aim to the provide an in-depth understanding of this \texttt{R} package and explore its functionality.

@@

A set of attributes is conceptualised as a space (termed an "attribute space"). Each planning unit may be thought to exist as a single point inside a given space. For example, a decision maker may require a prioritisation that represents populations along climatic gradients. To achieve this, the decision maker might use an "climatic" attribute space with dimensions relating to mean annual temperature ($^{\circ}$C) and precipitation (mm). Any given combination of temperature and precipitation may be conceived as a point in this environmental space. By associating planning units with climatic data, they can be mapped from geographic space to this environmental attribute space. In addition to planning units, attribute spaces also contain demand points.

>>>>

4. The \\texttt{raptr R} package provides a unified framework for achieving spatial conservation prioritisations that secure an adequate and representative sample of features. Prioritisations that fulfill that , resulting in a greater chance of achieving long-term biodiversity persistence than area-based planning alone.

@@

Perhaps the most fundamental aim of conservation is to maximise the long-term persistence of biodiversity [@r423; @r255]. To achieve this, conservation actions must preserve biodiversity patterns (eg. populations, species, ecosystems), but also crucially the processes that sustain them. One of the major tangible achievements of modern conservation has been the act of setting aside areas for preservation [@r440]. Reserve networks buffer species from gross threatening processes and set the stage for enhanced management interventions [@r465]. Since the resources available for conservation action are limited, protected area networks must be sited in places that satisfy conservation objectives for minimal cost [@r255].

Today, the most widely used conservation planning tools focus on biodiversity patterns [\texttt{Zonation} and \texttt{Marxan}; @r429; @r22]. These tools allow decision makers to obtain solutions that secure a proportion of the geographic range of biological features (populations, species, or ecosystems of conservation interest). One method to incorporate data on the ecological and environmental processes that sustain features is to conserve a representative sample of the physical attributes that underpin these processes. To achieve this, current methods typically involve partitioning each feature into sub-groups based on an attribute variable that relates to a biodiversity process of interest [@r9; @r247; @r448]. For instance, by dividing species distributions into sub-groups according to habitat discontinuities, and ensuring that each sub-group is represented in the solution, conservation planners can obtain prioritisations that promote adaptive processes [@r2]. However, using this method is challenging because biodiversity data are often continuous and multi-dimensional.

Simply partitioning features based on a continuous variables and setting targets for each sub-group may not necessarily result in a solution that secures a representative sample of the continuous variable. For instance, if a feature was split into four sub-features based on a single continuous variable, and a solution preserve 10 % of the geographic range of each sub-feature, how much of the original continuous variation does the solution represent? The answer is not immediately clear. If the solution sites reserves evenly along the original variable, then the solution may sample far more than 10 % of the original variation. Whereas, if the solution sites reserves near the values used to divide the continuous values [as observed when partitioning continuous geographic variation based on political boundaries; @r483], then it may sample much less of the original variation. Conventional reserve selection problems cannot differentiate between either of these solutions--they are both optimal--though the poorer solution may be favored when using additional constraints to reduce fragmentation. This limitation has not arisen recently. In fact, it dates back to some of the earliest reserve selection methods [eg. @r295].

During the early 1990s, Faith and Walker [-@r480] developed a decision support tool (\texttt{DIVERSITY}) to identify prioritisations that secure a representative sample of continuous environmental variation across a study area. Although this method was originally pitched as an alternative to species-based conservation planning [@r481; @r218; @r203], it can also identify solutions that conserve a representative sample of variation across a single species' range. Recent work has built on this reserve selection method by solving problems with more advanced optimisation routines [@r482]. However, \texttt{DIVERSITY}-based reserve selection methods are not widely used today, in part, because they do not accommodate explicit amount-based targets for multiple features (_cf_. \texttt{Marxan} and \texttt{Zonation}).

One of the key issues in reserve selection is the lack of a unified decision support tool that uses explicit and separate targets to obtain solutions that secure an adequate amount of habitat for each feature and a representative sample of each feature. To begin to fill this gap, here we unite the ideas underpinning \texttt{DIVERSITY} and \texttt{Marxan} into new formulations of the reserve selection problem, and implement them in the \texttt{raptr R} package. This \texttt{R} package provides decision makers with the tools to generate prioritisations based on data that relate to biodiversity patterns and processes. Here, we aim to the provide an in-depth understanding of this \texttt{R} package and explore its functionality.

@@

A set of attributes is conceptualised as a space (termed an "attribute space"). Each planning unit may be thought to exist as a single point inside a given space (). For example, a decision maker may require a prioritisation that represents populations along climatic gradients. To achieve this, the decision maker might use an "climatic" attribute space with dimensions relating to mean annual temperature ($^{\circ}$C) and precipitation (mm). Any given combination of temperature and precipitation may be conceived as a point in this environmental space. By associating planning units with climatic data, they can be mapped from geographic space to this environmental attribute space. In addition to planning units, attribute spaces also contain demand points.
